### 2017-01-13 intro.tex
sir corrections first intro only

\chapter{Introduction}
Optimization problems can be classified into several branches depending upon the nature of the problem. Consider the following optimization problem which is to find a $\theta^{*}$ that minimizes the objective function $f(\theta)$:
%\begin{equation} 
\begin{align}\label{eq:opt}
 \underset{\theta \in \Theta}{\min} ~f(\theta),
\end{align}
%\end{equation}
where $f \colon \mathbb{R}^N \to \mathbb{R}$ is called the objective function, $\theta$ is tunable N-dimensional parameter and $\Theta \subseteq \mathbb{R}^N$ is the constraint set in which $\theta$ takes values. 
\section{Classification of optimization problems based on objective function}\label{sc:objclass}
If we have complete information about $f$ and its derivatives etc., and about the set $\Theta$ then \eqref{eq:opt} would be a \emph{deterministic optimization} problem. In the real world unfortunately, many problems do not fall in this class. However, Optimization problems involving uncertainties are very common in a variety of engineering disciplines such as transportation systems, manufacturing, networks, healthcare and finance.

The setting in \emph{stochastic optimization}, however presumes that we have little knowledge on the structure of $f$ and moreover $f$ cannot be obtained directly, but rather is an expectation of another quantity $h(\theta,\xi)$, to which we have access, i.e.,
\begin{align}\label{eq:stopt}
 f(\theta) \equiv E_{\xi}[h(\theta,\xi)], 
\end{align}
 where $\xi$ comprises the randomness in the system and one is allowed to observe only these $h(\theta,\xi)$ samples, largely  because one dose not have access to distribution of noise $\xi$. This kind of optimization problems are more challenging because of the added complexity of not knowing $f$ and to find $\theta^{*}$ only on the basis of aforementioned noisy samples. The large number of input variables and the lack of precise system model may prohibit analytical solution approaches and a viable alternative is to employ a simulation-based optimization approach. As illustrated in Figure \ref{fig:so}, the idea here is to simulate a few times the stochastic system under consideration until a good enough solution is obtained. A natural solution approach is to devise an algorithm that incrementally updates the parameter, say $\theta_n$, in the descent direction using the gradient and/or Hessian of the objective $f$. However, in practice, one can only obtain estimates of the function $f$ through black-box simulation and the challenge is to estimate the gradient and/or Hessian of $f$ from function samples. \par

@@

Suppose the objective function $f$ has the form $f(\theta) = \sum_{i=1}^{n} E[h_{i}(X_{i})]$ , where $n$ denotes the number of stages and $X_{i}$ is the state of the underlying process in stage $i$ and $h_{i}$ denotes a stage and state dependent cost function. Then this type of optimization problems are called \emph{multi-stage problems}. Let $\theta = (\theta_{1},\ldots,\theta_{n})^{T}$,  with each $\theta_{j}$ being scalar and Let $X_{i}$ depend on $\theta_{1},\ldots,\theta_{i}$. The idea is that optimization can be done one stage at a time over $n$ stages after observing the state $X_{i}$ in each stage $i$. The value $\theta_{i}$ in stage $i$ has bearing on the cost of all subsequent stages $i+1,...,n$. This is  a problem of dynamic optimization. Approaches such as Dynamic programming can be used to solve this optimization problem.\par
Now we consider sub-class of multi-stage problems having infinite number of stages and having \emph{long-run average cost function} as objective. Objective function here is of the form:
\begin{align}\label{eq:long}
 f(\theta) = \lim_{n \to \infty} \frac{1}{n} ~ E \left[ \sum_{i=1}^{n} h_{i}(X_{i})\right ] ,
\end{align}
where $X_{i}$ is defined same as before and we assume it depends on parameter $\theta$. Objective function \eqref{eq:long} in most cases is not known analytically. In such cases it makes difficult for the search procedure to update the search parameter with out estimating the cost over infinite long trajectory. Another subclassifications of optimization problems is based on feasible region $\Theta$, see \eqref{eq:opt}.

\section{Classification of optimization problems based on feasible region}
Consider the basic optimization problem in \eqref{eq:opt}. Optimization problems can be classified based on the structure of feasible region $\Theta$. Where set $\Theta$ can be of following types:

\begin{inparaenum}[\bfseries (i)]
 \item \emph{Discrete}
\item \emph{Continuous}
\item \emph{Hybrid} (continuous in some dimensions and discrete in others).
\end{inparaenum}
Hence corresponding optimization problems are called discrete, continuous, hybrid optimization problems. These subclassifications are common to both deterministic and stochastic optimization problems presented in section \ref{sc:objclass}.

\section{Optimization methods via simulation }\label{sc:dis}
Optimization via simulation is fast developing area for both research and practice. As mentioned before, for stochastic optimization problems simulation is a viable approach compared to solving them analytically. Multiple simulations replications must be performed in order to get good estimate of $E_{\xi}[h(\theta_i,\xi)]$, which is function value at $\theta_i$. The standard approach to estimate is by the sample mean
\begin{align}\label{eq:sm}
\overline f(\theta_i) = \frac{1}{n} ~  \sum_{j=1}^{n} h(\theta,\xi_j),
\end{align}
where $n$ is the number of simulation replications and $\xi_j$ is the j-th sample of randomness. As $n \to \infty$, $\overline f(\theta_i) \to E_{\xi}[h(\theta_i,\xi)]$. Primary concerns in stochastic optimization via simulation are 
\begin{inparaenum}[\bfseries (i)]
	\item $n$ must be large enough to get good estimate of  $E_{\xi}[h(\theta_i,\xi)]$, which has ultimate impact on the final solution that the optimization procedure need to find; and
	\item The above procedure must be applied for many different parameters ,i.e., $\theta$'s in order to find best $\theta$.
\end{inparaenum}
Hence one of the key criteria for comparing different simulation-based optimization approaches is by total number of simulations performed by them to attain specific function value. Note that different approaches may take different simulations per single iteration. In this section we present methods only for Continuous and Discrete optimization via simulation problems which often occur in practise. However, hybrid optimization problems are not frequently encountered. From  now on, we consider the setting of optimization problems mentioned in this section as the stochastic optimization setting presented in section \ref{sc:objclass}, i.e., when objective function is an expected value over certain noisy cost measurements. 
\subsection{Discrete optimization via simulation}
Optimization problems for which $\theta \in \Theta$ is discrete and integer ordered and $\Theta$ is a subset of $N$-dimensional integers are known as Discrete optimization via simulation problems. In the discrete case $\theta$ can take only countable (finite or infinite) number of values. Discrete optimization problems are some times called as \emph{combinatorial optimization} and have applications in resource allocation, network routing, policy planning etc.
In this section we consider the case of $\Theta$ being finite i.e., $\Theta = \{\theta_{1},\theta_{2},\ldots,\theta_{k}\}$ where $k \in \mathbb{N}$. Discrete optimization problems are divided into two sub-classes when feasible region is  either small (often less than 100 ) or large. we present optimization methods in literature separately for each class.\\
%\begin{enumerate}
\subsubsection{$\Theta$ is small}
In this case we can simulate all possible solutions and select the best among them. However, unlike in deterministic optimization, simulating only once for each parameter is not enough since the objective function is noisy in stochastic optimization setting. Hence the main question is how to conduct multiple simulation effectively for each parameter in order to determine best $\theta$.  Two popular methods in literature for this case are Ranking and Selection (R\&S)\cite{bech} and Multiple comparison procedures (MCP) \cite{mcp}. These methods are well suited for simulation since under lying assumptions such as normality and independence of observations can be met by careful selection procedures.
Traditional methods are applicable for maximum size of 20 while some recent techniques allow more. 
\begin{enumerate}

\item \textbf {Ranking and selection }
R\&S procedures are statistical procedures specially developed to select the best parameter or subset that contains the best parameter. Majority of work in R\&S is classified into  two categories. The first one is called \emph{Indifferent-zone ranking}\cite{bech,izrs} which is aimed  to select the best and another one is \emph{Subset selection}\cite{ssrs} where the procedures are designed to find subset consisting of best. Again there are two approaches for selecting the best problem : \emph{the frequentist approach} and \emph{ Bayesian approach}. One of the popular streams of Bayesian approach procedures is  optimal computing budget allocation (OCBA) procedures \cite{ocbabook,chen2000} in which simulation budget is allocated in a manner that maximises the posterior probability of correct selection. This method is widely considered the best procedure for small scale discrete optimization.

\item \textbf {MCPs}
Like R\&S, MCPs attempt to identify the best parameter. But MCPs approach optimization problem as a statistical inference problem and do not guarantee a decision. Three main classes of MCPs that are used in practise are \emph{Multiple Comparison approach} (MCA) , \emph{Multiple Comparison with the Best} (MCB) and \emph{Multiple Comparisons with Control} (MCC). The most popular among these is MCB. In particular, MCB looks at $f(\theta_{j}) - opt_{i \neq j} f(\theta_{i})$ for $j = 1,2,\ldots,k$ to determine that $j_{*}$ with $f(\theta_{j_{*}}) - opt_{ i \neq j_{*}} f(\theta_{i}) > 0$. Simultaneous confidence intervals $f(\theta_{j}) - opt_{i \neq j} f(\theta_{i})$ for $j = 1,2,\ldots,k$ can be used to determine $j_{*}$ by looking  for the confidence interval with lower confidence limit that is zero.
\end{enumerate}
For a comprehensive treatment of these methods one can refer \cite{swish,swish2}.
\subsubsection{$\Theta$ is large}
When $\Theta$ is large enumeration becomes too expensive to conduct. Some sort of search techniques like the ones in deterministic optimization must be applied to avoid simulating for all the parameters while ensuring a high chance of finding the best or good parameters. Some approaches in this category include the following : 
\begin{enumerate}
\item \emph{Model-based approaches:} Iterative algorithms using statistical methods search the feasible region to improve upon the candidate parameter. One can  use \emph{gradient/Hessian information} with respect to parameter to help the search similar to deterministic case. But due to lack of availability of function analytical form and distribution of noise one needs to estimate the gradient/Hessian from samples of function $f$. However gradient/Hessian  estimation could be  quite noisy due to stochastic nature involved  in these problems. There are several approaches  which are used in continuous case can also be used in this setting to estimate the gradient/Hessian when simulation model is treated as black box. We will present detailed literature survey of these approaches in sections \ref{sec:gradest} and \ref{sec:hesest}. 
\item \emph{Metaheuristic:} These are gradient free approaches. They include approaches such as genetic algorithms, evolutionary algorithms, simulated annealing \cite{simann}, tabu search \cite{tabu} and cross entropy \cite{crossent}. Most of these iterative algorithms start with an initial population of parameters and in each iteration elite parameters are selected from previous population and generates  better parameter population as search progresses. 
\end{enumerate}

@@

Optimization problems for which $\theta \in \Theta$ is a vector of continuous decision variables and $\Theta$ is a convex subset of $\mathcal{R}^N$ are called continuous optimization via simulation  problems. In these problems $\theta$ can take uncountable number of values. These kind of problems frequently occur in applications such as model fitting, adaptive control, neural network training, signal processing etc. A natural solution approach is to devise an algorithm that incrementally updates the parameter, say $\theta_n$, in the descent direction using the gradient and/or Hessian of the objective $f$. Stochastic approximation algorithms are best suited and most studied and used algorithms for solving continuous optimization problems via simulation.\par

@@

where $\theta_{n}$ is the solution found at iteration $n$. $ \widehat{\nabla} f(\theta_{n})$ is an estimate of the gradient $ {\nabla} f(\theta_{n})$ and $\{a_{n}\}$ is a sequence of positive reals satisfy following properties: $\sum_{n=1}^{\infty} a_{n}  = \infty$,  $\sum_{n = 1}^{\infty} a_{n}^{2} < \infty$, and $\Gamma_{\Theta}$ denotes the projection of a point outside $\Theta$ onto $\Theta$. Under appropriate conditions, as number of iterations goes to infinity one can guarantee convergence to the local minimum with probability one. The SA algorithm can be considered as a stochastic analog of gradient decent method which seeks the next solution along negative gradient direction. The main difference between the SA algorithm and steepest descent algorithm is even though gradient estimate in former algorithm is a noisy estimate it requires only weak assumptions (like $E \left[\widehat{\nabla} f(\theta_{n})\right] - \nabla f(\theta_{n}) \to 0 $ at a certain rate) for converging to local minimum. In practice, the performance of the SA algorithm is quite sensitive to the sequence $\{a_{n}\}$ for example in the case $ a_n = a/n$, the convergence is highly dependent on the choice of $a$. However, in practice, one can only obtain estimates of the function $f$ through black-box simulation and the challenge is to estimate the gradient and/or Hessian of $f$ from function samples.  In the following sections we will present a brief survey of existing methods to estimate gradient and Hessian from function samples.

@@

When the simulation model is treated as a block box, the traditional means of forming the estimate of gradient is by using finite-difference stochastic approximation (FDSA) method. There are mainly two variations in the FDSA scheme. First one is  two-sided gradient approximation which involve function measurements $f(\theta_n + \delta_n e_i)$ and $f(\theta_n - \delta_n e_i)$, where $e_i$ denotes $i$th column of a identity matrix of size $N \times N$. Denote these respective values by $y_{ni}^+$ and $y_{ni}^-$, i.e., 

@@

In the above, we assume the noise vector $(\xi_{ni}^+ - \xi_{ni}^-, i=1,2,\ldots,N)\tr$ is a martingale difference sequence for every $n \ge 0$, the sequence of the perturbation constants $\{\delta_n, n\ge 0\}$ is a positive and asymptotically vanishing sequence. The gradient estimate of objective function using this scheme is given by

@@

\right)

@@

Second one is  one-sided gradient approximations which involve function measurements $f(\theta_n+\delta_n e_i)$ and $f(\theta_n)$. Let us denote these values by  $y_{ni}^+$, $y_{n}$  respectively, i.e., 
$y_{ni}^+ = f(\theta_n+\delta_n e_i) + \xi_{ni}^+$, $y_{n} = f(\theta_n) + \xi_n$,  where we assume the noise terms $(\xi_{ni}^+ - \xi_{n}, i=1,2,\ldots,N)\tr$ satisfy the martingale difference sequence property for $n \ge 0$. The gradient estimate of objective function using this scheme is given by

@@

It is important to determine conditions under which $\theta_n$ converges to $\theta^*$, when one uses estimates shown in \eqref{eq:twofdsa} or \eqref{eq:onefdsa} used in \eqref{saalg}. The convergence theory for FDSA algorithm is similar to convergence theory of Robbins and Monro root finding SA algorithm \cite{rm}. However difficulties arise due to a bias in gradient approximation, i.e. $\E \left[\widehat{\nabla} f(\theta_{n})|\F_n \right] - \nabla f(\theta_{n}) $, where $\F_n = \sigma(x_m, m\le n)$ denotes the underlying sigma-field. The conditions required for showing convergence of this algorithm is as follows

@@

\item $f:\R^N\rightarrow \R$ is three-times continuously differentiable\footnote{Here $\nabla^3 f(\theta) = \dfrac{\partial^3 f (\theta)}{\partial \theta\tr \partial \theta\tr \partial \theta\tr}$ denotes the third derivate of $f$ at $\theta$ and $\nabla^3_{i_1 i_2 i_3} f(\theta)$ denotes the $(i_1 i_2 i_3)$th entry of $\nabla^3 f(\theta)$, for $i_1, i_2, i_3=1,\ldots, N$.}  with $\left|\nabla^3_{i_1 i_2 i_3} f(\theta) \right| < \alpha_0 < \infty$, for $i_1, i_2, i_3=1,\ldots, N$ and for all $\theta\in \R^N$. 

@@

Note that two-sided gradient estimation scheme requires $2N$ simulations of objective function in order to obtain single gradient estimate, where as one-sided gradient estimation scheme requires only $N+1$ simulations, where $N$ is the dimension of the vector $\theta$. Dependence on $N$ for number of simulations to get one gradient estimate makes this procedure highly computationally expensive when $N$ is large. Hence one requires gradient estimate procedure that is independent of dimension of $\theta$.

@@

Simultaneous perturbation (SP) methods are a popular and efficient approach for estimating gradient/Hessian from function samples, especially in high dimensional problems - see \cite{bhatnagar-book} for a comprehensive treatment of this subject matter. Simultaneous perturbation stochastic approximation (SPSA) is a popular SP method. The first-order SPSA algorithm, henceforth referred to as 1SPSA, was proposed in \cite{spall2005introduction}.  1SPSA scheme for approximating gradient requires only two simulations of function measurements irrespective of parameter dimension. The idea in this scheme is to perturb all components of parameter randomly for getting two function measurements. The function measurements required for this scheme corresponds to $\theta_n + \delta_n \Delta_n$ and $\theta_n - \delta_n \Delta_n$, respectively, where $\Delta_n = (\Delta_{n1},\ldots,\Delta_{nN})\tr$ is any vector consists of i.i.d, mean-zero, symmetric random variables whose inverse moments are bounded. The simplest and most commonly used perturbation distribution of random variables being used are symmetric Bernoulli distribution with $\Delta_{ni} = \pm 1 \,\,w.p.\,\,1/2$, $i=1,\ldots,N,\,\, n\ge 0$. The gradient estimate of objective function using this scheme is as follows

@@

The reason why 1SPSA algorithm is valid gradient estimation scheme can be seen easily from Taylor expansions of $f(\theta_n+\delta_n \Delta_n$ and $f(\theta_n-\delta_n \Delta_n)$.

@@

&f(\theta_n + \delta_n \Delta_n) = f(\theta_n) + \delta_n \Delta_n\tr \nabla f(\theta_n) + \frac{\delta_n^2}{2} \Delta_n\tr \nabla^2 f(\theta_n) \Delta_n + o(\delta_n^2)\\
&f(\theta_n - \delta_n \Delta_n) = f(\theta_n) - \delta_n \Delta_n\tr \nabla f(\theta_n) + \frac{\delta_n^2}{2} \Delta_n\tr \nabla^2 f(\theta_n) \Delta_n + o(\delta_n^2)

@@

From above two equations, the $i$th component of gradient approximation \eqref{eq:spsa} is given by

@@

\E\left[\frac{f(\theta_n + \delta_n \Delta_n) - \theta_n - \delta_n \Delta_n)}{2\delta_n \Delta_{ni}} | \F_n \right] = \nabla_i f(\theta_n) + o(\delta_n)

>>>>

\chapter{Introduction}
Optimization problems can be classified into various categories depending upon the nature of the problem. Consider the  problem of finding a $\theta^{*}$ that minimizes the objective function $f(\theta)$:
%\begin{equation} 
\begin{align}\label{eq:opt}
 \underset{\theta \in \Theta}{\min} ~f(\theta),
\end{align}
%\end{equation}
where $f \colon \mathbb{R}^N \to \mathbb{R}$ is called the objective function, $\theta$ is a tunable $N$-dimensional parameter and $\Theta \subseteq \mathbb{R}^N$ is the constraint set in which $\theta$ takes values. 
\section{Classification of optimization problems based on objective function}\label{sc:objclass}
If we have complete information about $f$ and its derivatives etc., and about the set $\Theta$ then \eqref{eq:opt} would be a \emph{deterministic optimization} problem. In the real world unfortunately, many problems do not fall in this class. However, optimization problems involving uncertainties are very common in a variety of engineering disciplines such as transportation systems, manufacturing, networks, healthcare and finance.

The setting in \emph{stochastic optimization}, however, presumes that we have little knowledge on the structure of $f$ and moreover $f$ cannot be obtained directly, but rather is an expectation of another quantity $h(\theta,\xi)$, to which we have access, i.e.,
\begin{align}\label{eq:stopt}
 f(\theta) \equiv E_{\xi}[h(\theta,\xi)], 
\end{align}
 where $\xi$ comprises the randomness in the system and one is allowed to observe only the $h(\theta,\xi)$ samples, largely  because one does not have access to the distribution of noise $\xi$. These kinds of optimization problems are more challenging because of the added complexity of not knowing $f$ and to find $\theta^{*}$ only on the basis of the aforementioned noisy samples. The large number of input variables and the lack of a precise system model may prohibit analytical solution approaches and a viable alternative is to employ a simulation-based optimization approach. As illustrated in Figure \ref{fig:so}, the idea here is to simulate a few times the stochastic system under consideration until a good enough solution is obtained. A natural solution approach is to devise an algorithm that incrementally updates the parameter, say $\theta_n$, in the descent direction using the gradient and/or Hessian of the objective $f$. However, in practice, one can only obtain estimates of the function $f$ through black-box simulation and the challenge is to estimate the gradient and/or Hessian of $f$ from function samples. \par

@@

Suppose the objective function $f$ has the form $f(\theta) = \sum_{i=1}^{n} E[h_{i}(X_{i})]$, where $n$ denotes the number of stages, $X_{i}$ is the state of the underlying process in stage $i$ and $h_{i}$ denotes a stage and state dependent cost function. Then such optimization problems are called \emph{multi-stage problems}. Let $\theta = (\theta_{1},\ldots,\theta_{n})^{T}$,  with each $\theta_{j}$ being a scalar and let $X_{i}$ depend on $\theta_{1},\ldots,\theta_{i}$. The idea is that optimization can be done one stage at a time over $n$ stages after observing the state $X_{i}$ in each stage $i$. The value $\theta_{i}$ in stage $i$ has a bearing on the cost of all subsequent stages $i+1,...,n$. This is  a problem of dynamic optimization. Approaches such as dynamic programming can be used to solve this optimization problem.\par
One may also consider a sub-class of multi-stage problems having an infinite number of stages and with the \emph{long-run average cost function} as objective. The objective function here is of the form:
\begin{align}\label{eq:long}
 f(\theta) = \lim_{n \to \infty} \frac{1}{n} ~ E \left[ \sum_{i=1}^{n} h_{i}(X_{i})\right ] ,
\end{align}
where $X_{i}$ is defined  as before and is a function of the parameter $\theta$. The objective function \eqref{eq:long} in most cases is not known analytically. In such cases it becomes difficult for the search procedure to update the search parameter without estimating the cost over an infinitly long trajectory. Such  objective functions are  common in reinforcement learning applications. Another sub-class of optimization problems are those based on the feasible region $\Theta$, see \eqref{eq:opt}.

\section{Classification of optimization problems based on constrained feasible region}
Consider the basic optimization problem in \eqref{eq:opt}. The problem depends critically on the structure of the feasible region $\Theta$, where $\Theta$ can be of the following types:

\begin{inparaenum}[\bfseries (i)]
 \item \emph{Discrete}
\item \emph{Continuous}
\item \emph{Hybrid} (continuous in some dimensions and discrete in others).
\end{inparaenum}
The corresponding optimization problems are called discrete, continuous or hybrid optimization problems. These classifications are common to both deterministic and stochastic optimization problems presented in section \ref{sc:objclass}.

\section{Optimization methods via simulation }\label{sc:dis}
Optimization via simulation is a fast developing area for both researchers and practioners. As mentioned before, for stochastic optimization problems, simulation is a viable approach compared. Multiple simulation replications must be performed in order to get a good estimate of $E_{\xi}[h(\theta_i,\xi)]$, which is the function value at parameter $\theta_i$. The standard approach to estimate this is via the sample mean
\begin{align}\label{eq:sm}
\overline {f_n} (\theta_i) = \frac{1}{n} ~  \sum_{j=1}^{n} h(\theta,\xi_j),
\end{align}
where $n$ is the number of simulation replications and $\xi_j$ is the $j$th sample of randomness. As $n \to \infty$, $\overline{ f_n} (\theta_i) \to E_{\xi}[h(\theta_i,\xi)]$. The primary concerns in stochastic optimization via simulation are 
\begin{inparaenum}[\bfseries (i)]
	\item $n$ must be large enough to get a good estimate of  $E_{\xi}[h(\theta_i,\xi)]$, which has an impact on the final solution that the optimization procedure needs to find; and
	\item The above procedure must be applied for many different parameters, i.e., $\theta$'s in order to find the best $\theta$.
\end{inparaenum}
Hence one of the key criteria for comparing different simulation-based optimization approaches is through the total number of simulations performed by them to attain a specific function value. Note that different approaches may require different number of simulations per  iteration. In this section we present methods only for Continuous and Discrete optimization via simulation problems which often occur in practise. However, hybrid optimization problems are not frequently encountered. From  now on, we consider the setting of optimization problems mentioned in this section as the stochastic optimization setting presented in section \ref{sc:objclass}, i.e., when objective function is an expected value over certain noisy cost measurements. 
\subsection{Discrete optimization via simulation}
Optimization problems for which $\Theta$ is a discrete set are known as discrete optimization via simulation problems. In the discrete case $\theta$, can take only countable (finite or infinite) number of values. Discrete optimization problems are some times take the form of \emph{combinatorial optimization} problems and they have applications in resource allocation, network routing, policy planning etc.
In this section we consider the case of $\Theta$ being finite i.e., $\Theta = \{\theta_{1},\theta_{2},\ldots,\theta_{k}\}$ where $k \in \mathbb{N}$. Discrete optimization problems are divided into two sub-classes when the feasible region is  either small (often less than 100) or is large. We present optimization methods in the literature separately for each class.\\
%\begin{enumerate}
\subsubsection{$\Theta$ is small}
In this case we can simulate all possible solutions and select the best among them. However, unlike in deterministic optimization, simulating the system only once for each parameter is not enough since the objective function is noisy in the stochastic optimization setting. Hence the main question is how to conduct multiple simulations effectively for each parameter in order to determine the  best parameter $\theta$.  Two popular methods in literature for this case are Ranking and Selection (R\&S)\cite{bech} and Multiple comparison procedures (MCP) \cite{mcp}. These methods are well suited for simulation since  the underlying assumptions such as normality and independence of observations can be met through careful selection procedures.
Traditional methods in this setting are applicable for a maximum set size of 20 while some recent techniques allow more. 
\begin{enumerate}

\item \textbf {Ranking and selection }
R\&S procedures are statistical procedures specially developed to select the best parameter or subset that contains the best parameter. Majority of the work in R\&S is classified into  two categories. The first  is called \emph{Indifferent-zone ranking}\cite{bech,izrs} which is aimed  to select the best parameter  and the other one is \emph{Subset selection}\cite{ssrs} where the procedures are designed to find a subset of parameters consisting of the best parameter. Again there are two approaches for selecting the best parameter : \emph{the frequentist approach} and \emph{ Bayesian approach}. One of the popular approaches in the  Bayesian stream is the optimal computing budget allocation (OCBA) procedure \cite{ocbabook,chen2000} where the simulation budget is allocated in a manner that maximizes the posterior probability of correct selection. This method is widely considered the best procedure for small scale discrete optimization.

\item \textbf {MCP}
Like R\&S, MCP attempts to identify the best parameter. But MCPs approach the optimization problem as a statistical inference problem and do not guarantee a decision. Three main classes of MCP that are used in practise are \emph{Multiple Comparison approach} (MCA) , \emph{Multiple Comparison with the Best} (MCB) and \emph{Multiple Comparisons with Control} (MCC). The most popular among these is MCB. In particular, MCB looks at $f(\theta_{j}) - opt_{i \neq j} f(\theta_{i})$ for $j = 1,2,\ldots,k$ to determine  $j_{*}$ such that $f(\theta_{j_{*}}) - opt_{ i \neq j_{*}} f(\theta_{i}) > 0$. Simultaneous confidence intervals $f(\theta_{j}) - opt_{i \neq j} f(\theta_{i})$ for $j = 1,2,\ldots,k$ can be used to determine $j_{*}$ by looking  for the confidence interval with the lower confidence limit of zero.
\end{enumerate}
For a comprehensive treatment of these methods, see \cite{swish,swish2}.
\subsubsection{$\Theta$ is large}
When $\Theta$ is large, simulating for each parameter value is expensive. Some sort of search techniques like the ones for deterministic optimization must be applied to avoid simulating for all the parameters while ensuring a high chance of finding the best or good parameters. Some approaches in this category include the following : 
\begin{enumerate}
\item \emph{Model-based approaches:} Iterative algorithms using statistical methods search the feasible region to improve upon the candidate parameter. One can  use \emph{gradient/Hessian information} with respect to the parameter to help the search similar to deterministic case. But due to lack of knowledge of the function in analytical form and distribution of noise one needs to estimate the gradient/Hessian from samples of the function $f$. However, gradient/Hessian  estimation could be  quite noisy due to the stochastic nature involved  in these problems. Various approaches  that are used in the continuous case can also be used in this setting to estimate the gradient/Hessian when the simulation model is treated as a black box. We will present detailed literature survey of these approaches in sections \ref{sec:gradest} and \ref{sec:hesest}. 
\item \emph{Metaheuristic:} These are gradient free approaches. They include approaches such as genetic algorithms, evolutionary algorithms, simulated annealing \cite{simann}, tabu search \cite{tabu} and cross entropy \cite{crossent}. Most of these iterative algorithms start with an initial population of parameters and in each iteration elite parameters are selected from the previous population and a better parameter population is generated as search progresses. 
\end{enumerate}

@@

Optimization problems for which $\theta \in \Theta$ is a vector of continuous decision variables and $\Theta$ is a convex subset of $\mathcal{R}^N$ are called continuous optimization via simulation  problems. In these problems $\theta$  takes an uncountable number of values. These kinds of problems frequently occur in applications such as model fitting, adaptive control, neural network training, signal processing etc. A natural solution approach is to devise an algorithm that incrementally updates the parameter, say $\theta_n$, in the descent direction using the gradient and/or Hessian of the objective $f$. Stochastic approximation algorithms are best suited and most studied and used algorithms for solving continuous optimization problems via simulation.\par

@@

where $\theta_{n}$ is the solution found at iteration $n$. Also, $ \widehat{\nabla} f(\theta_{n})$ is an estimate of the gradient $ {\nabla} f(\theta_{n})$ and $\{a_{n}\}$ is a sequence of positive reals satisfying the following properties: $\sum_{n=1}^{\infty} a_{n}  = \infty$,  $\sum_{n = 1}^{\infty} a_{n}^{2} < \infty$, and $\Gamma_{\Theta}$ denotes the projection operator. Under appropriate conditions, as the number of iterations goes to infinity one can guarantee convergence to the local minimum with probability one. The SA algorithm can be considered as a stochastic analog of the gradient decent method which seeks the next solution along the negative gradient direction. The main difference between the SA algorithm and steepest descent algorithm is even though the gradient estimate in the former algorithm is a noisy estimate it requires only weak assumptions (like $E \left[\widehat{\nabla} f(\theta_{n})\right] - \nabla f(\theta_{n}) \to 0 $ at a certain rate) for convergence to local minimum. In practice, the performance of the SA algorithm is quite sensitive to the sequence $\{a_{n}\}$.  For example in the case of $ a_n = a/n$, the convergence is highly dependent on the choice of $a$. However, in practice, one can only obtain estimates of the function $f$ through black-box simulation and the challenge is to estimate the gradient and/or Hessian of $f$ through function samples.  In the following sections we will present a brief survey of the existing methods to estimate the gradient and Hessian from function samples.

@@

When the simulation model is treated as a block box, the traditional means of forming the estimate of gradient is by using the finite-difference stochastic approximation (FDSA) method. There are mainly two variations in the FDSA scheme. First one is  two-sided gradient approximation which involves function measurements $f(\theta_n + \delta_n e_i)$ and $f(\theta_n - \delta_n e_i)$, where $e_i$ denotes $i$th column of the identity matrix of size $N \times N$. Denote these respective values by $y_{ni}^+$ and $y_{ni}^-$, i.e., 

@@

In the above, we assume the noise vector $(\xi_{ni}^+ - \xi_{ni}^-, i=1,2,\ldots,N)\tr$ is a martingale difference sequence for every $n \ge 0$, the sequence of the perturbation constants $\{\delta_n, n\ge 0\}$ is a positive and asymptotically vanishing sequence. The gradient estimate of the objective function using this scheme is given by

@@

\right).

@@

The second approach is of one-sided gradient approximation  involves function measurements $f(\theta_n+\delta_n e_i)$ and $f(\theta_n)$. Let us denote these values by  $y_{ni}^+$, $y_{n}$  respectively, i.e., 
$y_{ni}^+ = f(\theta_n+\delta_n e_i) + \xi_{ni}^+$, $y_{n} = f(\theta_n) + \xi_n$,  where we assume the noise terms $(\xi_{ni}^+ - \xi_{n}, i=1,2,\ldots,N)\tr$ satisfy the martingale difference sequence property for $n \ge 0$. The gradient estimate of the objective function using this scheme is given by

@@

It is important to determine conditions under which $\theta_n$ converges to $\theta^*$, when one uses estimates shown in \eqref{eq:twofdsa} or \eqref{eq:onefdsa}. The convergence theory for FDSA algorithm is similar to the convergence theory of Robbins and Monro root finding SA algorithm \cite{rm}. However, difficulties arise due to a bias in gradient approximation, i.e, $\E \left[\widehat{\nabla} f(\theta_{n})|\F_n \right] - \nabla f(\theta_{n}) $, where $\F_n = \sigma(\theta_m, m\le n)$ denotes the underlying sigma-field. The conditions required for showing convergence of this algorithm are as follows:

@@

\item $f:\R^N\rightarrow \R$ is three-times continuously differentiable\footnote{Here $\nabla^3 f(\theta) = \dfrac{\partial^3 f (\theta)}{\partial \theta\tr \partial \theta\tr \partial \theta\tr}$ denotes the third derivate of $f$ at $\theta$ and $\nabla^3_{i_1 i_2 i_3} f(\theta)$ denotes the $(i_1 ,i_2, i_3)$th entry of $\nabla^3 f(\theta)$, for $i_1, i_2, i_3=1,\ldots, N$.}  with $\left|\nabla^3_{i_1 i_2 i_3} f(\theta) \right| < \alpha_0 < \infty$, for $i_1, i_2, i_3=1,\ldots, N$ and for all $\theta\in \R^N$. 

@@

Note that the  two-sided gradient estimation scheme requires $2N$ simulations of the objective function in order to obtain a single gradient estimate, whereas the one-sided gradient estimation scheme requires only $N+1$ simulations, where $N$ is the dimension of the vector $\theta$. A requirement of number of simulations to be proportional to $N$ in order to get one gradient estimate makes this procedure highly computationally expensive when $N$ is large. Hence one requires a gradient estimation procedure that is independent of the dimension of $\theta$.

@@

Simultaneous perturbation (SP) methods are a popular and efficient approach for estimating gradient/Hessian from function samples, especially in high dimensional problems - see \cite{bhatnagar-book} for a comprehensive treatment of this subject matter. Simultaneous perturbation stochastic approximation (SPSA) is a popular SP method. The first-order SPSA algorithm, henceforth referred to as 1SPSA, was proposed in \cite{spall2005introduction}. The 1SPSA scheme for approximating gradient requires only two simulations or function measurements irrespective of the parameter dimension. The idea in this scheme is to use two function measurements by perturbing all components of the parameter randomly. The function measurements required for this scheme correspond to $\theta_n + \delta_n \Delta_n$ and $\theta_n - \delta_n \Delta_n$, respectively, where $\Delta_n = (\Delta_{n1},\ldots,\Delta_{nN})\tr$ is any vector consisting of i.i.d, mean-zero, symmetric random variables whose inverse moments are bounded. The simplest and most commonly used perturbation distribution of random variables being used are the symmetric Bernoulli random variables $\Delta_{ni} = \pm 1 \,\,w.p.\,\,1/2$, $i=1,\ldots,N,\,\, n\ge 0$. The gradient estimate of objective function using this scheme is as follows:

@@

The reason why 1SPSA algorithm is a valid gradient estimation scheme can be seen easily from Taylor expansions of $f(\theta_n+\delta_n \Delta_n$ and $f(\theta_n-\delta_n \Delta_n)$ as under:

@@

&f(\theta_n + \delta_n \Delta_n) = f(\theta_n) + \delta_n \Delta_n\tr \nabla f(\theta_n) + \frac{\delta_n^2}{2} \Delta_n\tr \nabla^2 f(\theta_n) \Delta_n + o(\delta_n^2),\\
&f(\theta_n - \delta_n \Delta_n) = f(\theta_n) - \delta_n \Delta_n\tr \nabla f(\theta_n) + \frac{\delta_n^2}{2} \Delta_n\tr \nabla^2 f(\theta_n) \Delta_n + o(\delta_n^2).

@@

From the above two equations, the $i$th component of the gradient approximation \eqref{eq:spsa} is given by

@@

\E\left[\frac{f(\theta_n + \delta_n \Delta_n) - f(\theta_n - \delta_n \Delta_n)}{2\delta_n \Delta_{ni}} | \F_n \right] = \nabla_i f(\theta_n) + o(\delta_n).
